An Analysis of daily shelter and overnight service usage in Toronto focusing on shelter occupancy and capacity

### 2024-09-27 paper/paper.qmd
Final Revision

---
title: "An Analysis of daily shelter and overnight service usage in Toronto focusing on shelter occupancy and capacity"
author: 
  - "Angel Xu"
thanks: "Code and data are available at: https://github.com/Anjojoo/Toronto-Overnight-Servive-Occupancy-Capacity."
date: today
date-format: long
abstract: "This paper analyzes the overnight service occupancy and capacity of shelters in Toronto, utilizing data from OpenDataToronto. Employing the R programming language along with various packages, the analysis focuses on shelter usage trends, specifically examining changes in occupancy rates and actual capacity over the years. The results indicate an overall upward trend in shelter occupancy rates, while a significant decline in actual capacity was observed in 2024. These findings highlight the need for strategic resource allocation and improved response strategies concerning shelter occupancy in the future."
format: 
  pdf: 
    toc: true
    toc-depth: 2
number-sections: true
bibliography: references.bib
---

@@

This paper's findings are crucial for policy makers, social service organizations, and urban planners. By shedding light on the trends in shelter occupancy and capacity, the study identifies gaps in service and potential areas for improvement. The analysis has implications for future resource allocation, as understanding the patterns of shelter usage can help design better support strategies, optimize capacity, and improve the overall response to homelessness in urban areas. Additionally, it can serve as a model for analyzing shelter systems in other cities, offering a framework for data-driven decision-making.

The remainder of this paper is structured as follows. Section 2(@sec-data) provides details on the data source, including the specific variables used, and describes the data cleaning process to prepare for analysis. This data section also outlines the tools and methodologies applied. Section 3(@sec-results) presents the key results, focusing on occupancy rates, service user counts, and actual capacity over time. Visualizations are used to intuitively convey these findings. Section 4{@sec-dis} offers a discussion of the results and their implications, summarizing conclusions and emphasizing the importance of balancing shelter occupancy and capacity for effective service provision. The comprehensive structure ensures that the reader gains a full understanding of what was analyzed, the findings, and their significance.

@@

## Data Measurement and Ethics

>>>>

---
title: "An Analysis of daily shelter and overnight service usage in Toronto focusing on shelter occupancy and capacity"
author: 
  - "Victoria Zheng"
thanks: "Code and data are available at: https://github.com/Anjojoo/Toronto-Overnight-Servive-Occupancy-Capacity."
date: today
date-format: long
abstract: "This paper analyzes the overnight service occupancy and capacity of shelters in Toronto, utilizing data from OpenDataToronto. Employing the R programming language along with various packages, the analysis focuses on shelter usage trends, specifically examining changes in occupancy rates and actual capacity over the years. The results indicate an overall upward trend in shelter occupancy rates, while a significant decline in actual capacity was observed in 2024. These findings highlight the need for strategic resource allocation and improved response strategies concerning shelter occupancy in the future."
format: 
  pdf: 
    toc: true
    toc-depth: 2
number-sections: true
bibliography: references.bib
---

@@

This paper's findings can be helpful for policy makers, social service organizations, and urban planners. By shedding light on the trends in shelter occupancy and capacity, the study identifies gaps in service and potential areas for improvement. The analysis has implications for future resource allocation, as understanding the patterns of shelter usage can help design better support strategies, optimize capacity, and improve the overall response to homelessness in urban areas. Additionally, it can serve as a model for analyzing shelter systems in other cities, offering a framework for global cases.

The remainder of this paper is structured as follows. @sec-data provides details on the data source, including the specific variables used, and describes the data cleaning process to prepare for analysis. Data tools and methodologies applied in this paper are also stated in this section. Measurement and ethics are considered as well. @sec-results presents the key results, focusing on occupancy rates, service user counts, and actual capacity over time; Visualizations are used to intuitively convey these findings. @sec-dis offers a discussion of the results and their implications,also emphasizing the importance of balancing shelter occupancy and capacity for effective service provision while limitations point out some weakness in this paper. The comprehensive structure ensures that the reader gains a full understanding of what was analyzed, the findings, and their significance.

@@

## Data Measurement and Ethics
For the data measurement of "Daily Shelter & Overnight Service Occupancy & Capacity" from opendatatoronto(@citeopendatatoronto), variables are mainly supported by Shelter Management Information System(SMIS). The occupancy data is recorded at 4:00 a.m. each morning, reflecting the previous evening's overnight service. Service user counts measures the number of service users at each program referring to the occupancy date. Actual capacity of beds and rooms are the data available on SMIS and used to calculate the occupancy rate with another available variable, occupied capacity which is also available on SMIS.

The ethical consideration included ensuring that the data used reflects accurate and timely reporting from the open-source platform and respecting the privacy of the individuals indirectly represented in the data. The data are collected anonymized and does not contain personally identifiable information, ensuring privacy and confidentiality of shelter users.
